\section{Generic Store}\label{chap:generic-store}

\newcommand{\angkv}{\langle k, v \rangle}
\newcommand{\angkvp}{\langle k', v' \rangle}

From now on, we say \textit{generic store} or simply \textit{store}, when talking about the data model we implemented in Coq, and \textit{context} when we refer to its role in the type system. The store can be used for purposes other than a typing context.

\begin{definition}
The generic stores are denoted by $S$ in this chapter, which are unordered lists of keys $k$ and values $v$, defined inductively as follows:
\begin{equation*}
S ::= \emptyset~|~S :: \angkv.
\end{equation*}
The empty store is denoted by $\emptyset$ and $::$ is the append operator.
\end{definition}

We would need decidablity of equality on the keys and values of the store. So, we assume the keys are chosen from a class $\mathbf{K}$ equipped with a reflexive relation $\asymp$ with the following property:
\begin{itemize}
    \item If $k_1 \asymp k_2$, then $k_1 = k_2$.
\end{itemize}
The axiom of extentionality is also required for the class of values, $\mathbf{V}$.

% \begin{listing}[t]
% \begin{minted}[
% frame=lines,
% framesep=2mm,
% baselinestretch=1.2,
% fontsize=\footnotesize,
% linenos
% ]{coq}
% Module Type KeyModuleType.

%   Parameter T : Type.
%   Parameter equal : T -> T -> bool.
%   Parameter eq_refl : forall x, equal x x = true.
%   Parameter eq_extensionality : forall x y, equal x y = true -> x = y.

% End KeyModuleType.

% Module Type ValModuleType.

%   Parameter T : Type.
%   Parameter equal : T -> T -> bool.
%   Parameter eq_refl : forall x, equal x x = true.

% End ValModuleType.
% \end{minted}
% \caption{Definition of the classes $\mathbf{K}$ and $\mathbf{V}$ in Coq}
% \label{listing:kvmoduletype}
% \end{listing}

\begin{definition}
Concatenation of two stores are defined as follows:

\begin{equation*}
S \circ S' =
\begin{cases}
    S, & S' = \emptyset, \\
    (S \circ S'') :: \angkv, & S' = S'' :: \angkv.
\end{cases}
\end{equation*}
\end{definition}

\begin{lemma}[Associativity]
For stores $S_1, S_2, S_3$, we have $(S_1 \circ S_2) \circ S_3 = S_1 \circ (S_2 \circ S_3)$.
\end{lemma}
\begin{proof}
By induction on $S_2$.
\end{proof}

\begin{definition}[Membership]
\label{defn:generic-context-membership}
By definition, stores are lists of ordered pairs. Membership is defined recursively as follows:
\begin{itemize}
    \item If $S = S' :: \angkv$, then $\angkv \in S$,
    \item If $\angkvp \in S'$ and $k \not= k'$, then $\angkvp \in S :: \angkv$. \qedhere
\end{itemize}
\end{definition}

The membership is defined so, so that appending a pair with the same key and a different value would overwrite the old record, in a sense:
\begin{itemize}
    \item $\angkv \in \emptyset :: \angkv$,
    \item $\angkv \not\in \emptyset :: \angkv :: \ang{k, v'}$.
\end{itemize}

\begin{definition}
A key $k$ is said to be a key in the store $S$, denoted as $\ang{k, \cdot} \in S$, iff for some value $v$ we have $\angkv \in S$.
\end{definition}

\begin{lemma}
Let $B$ and $C$ be stores, $k$ a key, and $v$ a value. Then:
\begin{itemize}
    \item If $\angkv \in C$, then $\angkv \in B \circ C$,
    \item If $\angkv \in B$ and $\ang{k, \cdot} \not\in C$, then $\angkv \in B \circ C$,
    \item If $\angkv \in B \circ C$, then either $\angkv \in B$ or $\angkv \in C$.
\end{itemize}
\end{lemma}

\begin{definition}[Equivalence]
We say the store $S$ is a subset of the store $S'$, writing $S \sqsubseteq S'$, iff for all $\angkv \in S$ we have $\angkv \in S'$. We say the two stores are equivalent, writing $S \equiv S'$, when $S \sqsubseteq S'$ and $S' \sqsubseteq S$.
\end{definition}

\begin{lemma}
The subset relation $\sqsubseteq$ on the stores is reflexive and transitive. Furthermore, the equivalence relation $\equiv$ is reflexive, transitive, and symmetric.
\end{lemma}

\begin{definition}
We say a store $S$ is \textit{duplicated}, if it has duplicated keys, recursively defined as follows:
\begin{itemize}
    \item If $S = S' :: \angkv$ and $\ang{k, \cdot} \in S'$, then $S$ is duplicated,
    \item If $S = S' :: \angkv$ and $S'$ is duplicated, so is $S$.
\end{itemize}
If $S$ is not duplicated, we call it \textit{unique-paired}.
\end{definition}

From now on, we would like to call two stores \textit{equal} if they are \textit{equivalent}.

\begin{axiom}[Extensionality]
\label{axiom:extentionality}
Let $B$ and $C$ be stores. Then $B = C$ if $B \equiv C$.
\end{axiom}

\begin{corollary}
Let $B$ and $C$ be stores. Then $B = C$ iff $B \equiv C$.
\end{corollary}
\begin{proof}
Follows from reflexiveness of $\equiv$.
\end{proof}

\begin{theorem}[Commute]
\label{theorem:commute}
Let $B$ and $C$ be stores. If $B \circ C$ is unique-paired, then $B \circ C = C \circ B$.
\end{theorem}

\begin{corollary}[Exchange]
\label{corollary:exchange}
For stores $A, B, C, D$, if $B \circ C$ is unique-paired, then $A \circ B \circ C \circ D = A \circ C \circ B \circ D$.
\end{corollary}

Corollary \ref{corollary:exchange} shows that the store we implemented can only be used with type systems that support the structural rule of exchange (hence, all except ordered ones).

\begin{lemma}[Append Commute]
\label{lemma:append_commut}
For any store $S$, keys $k, k'$, and values $v, v'$, if $k \not= k'$, then $S :: \ang{k, v} :: \ang{k', v'} = S :: \ang{k', v'} :: \ang{k, v}$.
\end{lemma}
\begin{proof}
By writing $S :: \ang{k, v} :: \ang{k', v'}$ as $S \circ (\emptyset :: \ang{k, v}) \circ (\emptyset :: \ang{k', v'})$, realizing that $(\emptyset :: \ang{k, v}) \circ (\emptyset :: \ang{k', v'})$ is unique-paired, and then using Corollary \ref{corollary:exchange} (exchange).
\end{proof}

There are many more lemmas and propositions stated and proved in the Coq implementation. One can find them in \texttt{ListCtx.v}.