\section{Declarative Typing}

\begin{listing}[t]
\begin{minted}[
frame=lines,
framesep=2mm,
baselinestretch=1.2,
fontsize=\footnotesize,
linenos,
escapeinside=<>,
mathescape=true
]{coq}
Module DeclarativeTyping<$\T$>.

Reserved Notation "G '⊢' t '|' T" (at level 60).

Inductive type : <$\mathfrak{G}$> -> <$\mathfrak{t}$> -> <$\T$> -> Prop :=
  | Var : forall <$\Gamma$> x T, (m<$\mathfrak{G}$>.append <$\Gamma$> x T) <$\vdash$> (<$\mathfrak{t}$>id x) | T
  | Unit_I : forall <$\Gamma$>, <$\Gamma$> <$\vdash$> <$\mathfrak{t}$>hole | Unit
  | Void_E : forall <$\Gamma$> t T, <$\Gamma$> <$\vdash$> t | Void -> <$\Gamma$> <$\vdash$> <$\mathfrak{t}$>holecase t | T
  | <$\T$>mult_I : forall <$\Gamma$> t1 t2 T1 T2, <$\Gamma$> <$\vdash$> t1 | T1 -> <$\Gamma$> <$\vdash$> t2 | T2 -> 
    <$\Gamma$> <$\vdash$> <$\mathfrak{t}$>pair t1 t2 | <$\T$>mult T1 T2
  | <$\T$>mult_E1 : forall <$\Gamma$> t T1 T2, <$\Gamma$> <$\vdash$> t | <$\T$>mult T1 T2 -> <$\Gamma$> <$\vdash$> <$\mathfrak{t}$>prj 1 t | T1
  | <$\T$>mult_E2 : forall <$\Gamma$> t T1 T2, <$\Gamma$> <$\vdash$> t | <$\T$>mult T1 T2 -> <$\Gamma$> <$\vdash$> <$\mathfrak{t}$>prj 2 t | T2
  | <$\T$>plus_I1 : forall <$\Gamma$> t T1 T2, <$\Gamma$> <$\vdash$> t | T1 -> <$\Gamma$> <$\vdash$> <$\mathfrak{t}$>inj 1 t | <$\T$>plus T1 T2
  | <$\T$>plus_I2 : forall <$\Gamma$> t T1 T2, <$\Gamma$> <$\vdash$> t | T2 -> <$\Gamma$> <$\vdash$> <$\mathfrak{t}$>inj 2 t | <$\T$>plus T1 T2
  | <$\T$>plus_E : forall <$\Gamma$> t t1 t2 x1 x2 T1 T2 S, <$\Gamma$> <$\vdash$> t | <$\T$>plus T1 T2 ->
    (m<$\mathfrak{G}$>.append <$\Gamma$> x1 T1) <$\vdash$> t1 | S -> (m<$\mathfrak{G}$>.append <$\Gamma$> x2 T2) <$\vdash$> t2 | S -> 
    <$\Gamma$> <$\vdash$> <$\mathfrak{t}$>case t x1 t1 x2 t2 | S
  | <$\T$>impl_I : forall <$\Gamma$> t x T S, (m<$\mathfrak{G}$>.append <$\Gamma$> x T) <$\vdash$> t | S -> 
    <$\Gamma$> <$\vdash$> <$\mathfrak{t}$>lambda x t | <$\T$>impl T S
  | <$\T$>impl_E : forall <$\Gamma$> t1 t2 T S, <$\Gamma$> <$\vdash$> t1 | <$\T$>impl T S -> <$\Gamma$> <$\vdash$> t2 | T -> 
    <$\Gamma$> <$\vdash$> <$\mathfrak{t}$>app t1 t2 | S
where "G '⊢' t '|' T" := (type G t T).

End DeclarativeTyping<$\T$>.
\end{minted}
\caption{Implementation of non-linear typing rules}
\label{listing:coq-declarativetypingT}
\end{listing}

In this section, we are going to describe declarative typing of the type system implemented in Coq.
%
Coq types $\t$ and $\e$ represent terms and expressions, based on \figref{fig:syntax-of-types-terms-expressions}.
Also,
\begin{itemize}
    \item \texttt{DeclarativeTyping}$\T$ is the implementation of typing rules for terms, based on \figref{fig:typing-rules-for-nonlinear-types} and represented in Listing \ref{listing:coq-declarativetypingT},
    \item \texttt{DeclarativeTyping}$\A$ is the implementation of typing rules for expressions, based on \figref{fig:typing-rules-for-linear-types},
    \item \texttt{DeclarativeTyping}$\T\A$ is the implementation of typing rules featuring lift and lower, to move between non-linear and linear types, based on \figref{fig:typing-rules-for-linear-and-nonlinear-types}.
\end{itemize}

The first obvious result of the implementation follows:

\begin{lemma}
If $\ang{x, \T} \in \Gamma$, then $\Gamma \vdash x : \T$.
\end{lemma}
\begin{proof}
By induction on the definition of membership (Definition \ref{defn:generic-context-membership}), either one of the following statements are true:
\begin{itemize}
    \item $\Gamma = \Gamma' :: \ang{x, \T}$, in which case the statement follows from the \texttt{Var} rule of \listref{listing:coq-declarativetypingT}.
    \item $\Gamma = \Gamma' :: \ang{x', \T'}$ and $\ang{x, \T} \in \Gamma'$ and $x \not= x'$. In this case, by induction we have $\Gamma' \vdash x : \T$, which means $\Gamma' = \Gamma'' :: \ang{x, \T}$. Using the append commute lemma (\lemref{lemma:append_commut}), we have
    \[
    \Gamma = \Gamma'' :: \ang{x, \T} :: \ang{x', \T'} = \Gamma'' :: \ang{x', \T'} :: \ang{x, \T},
    \]
    and so $\Gamma \vdash x : \T$ follows once more from \texttt{Var}. \qedhere
\end{itemize}
\end{proof}

The Coq implementation of this lemma is much less verbose:

\begin{minted}[escapeinside=<>,mathescape=true]{coq}
Lemma contains_<$\T$>_var : forall <$\Gamma$> x T,
  m<$\mathfrak{G}$>.contains <$\Gamma$> x T -> <$\Gamma$> <$\vdash$> <$\t$>id x | T.
Proof.
  intros. induction H.
  - subst s. apply T<$\T$>.Var.
  - inversion IHcontains. rewrite -> m<$\mathfrak{G}$>.append_commut_weak.
    apply T<$\T$>.Var. auto.
Qed.
\end{minted}