\section{Operational Semantics}

The operational semantics implementation is based on \figref{fig:operational-semantics-step}. Two immediate results are the following:

\begin{lemma}[$\t$-Preservation]
\label{lemma:t-preservation}
For all $\Gamma, \t, \t', \T$, if $\Gamma \vdash \t : \T$ and $\t \leadsto \t'$, then $\Gamma \vdash \t' : \T$.
\end{lemma}

The proofs to this lemma requires a simple rule on typing substitutions:
\[
\infer[\mbox{$\t$-Subst}]{\Gamma \vdash \t[x \mapsto \t'] : \T}{
    \Gamma, x : \T' \vdash \t : \T
    &
    \Gamma \vdash \t' : \T'
}
\]

Or as one might express in Coq:
\begin{minted}[escapeinside=<>,mathescape=true]{coq}
Axiom <$\t$>subst_type : forall <$\Gamma$> t t' T T' x,
  type<$\T$> (m<$\mathfrak{G}$>.append <$\Gamma$> x T') t T -> 
  type<$\T$> <$\Gamma$> t' T' -> 
  type<$\T$> <$\Gamma$> (SemanticsBase.<$\t$>subst t x t') T.
\end{minted}

Using this rule, the proof of Lemma \ref{lemma:t-preservation} is straightforward:

\begin{minted}[escapeinside=<>,mathescape=true]{coq}
Lemma <$\t$>_preservation : forall <$\Gamma$> t t' T, 
  type<$\T$> <$\Gamma$> t T -> <$\t$>sem t t' -> type<$\T$> <$\Gamma$> t' T.
Proof.
  intros. induction H0; inversion H.
  - inversion H2. auto.
  - inversion H2. auto.
  - eapply <$\t$>subst_type. inversion H3.  apply H8. apply H5.
  - eapply <$\t$>subst_type. apply H8. inversion H7. auto.
  - eapply <$\t$>subst_type. apply H9. inversion H7. auto.
Qed.
\end{minted}

A weaker but similar result is expressible for the linear part:

\begin{lemma}[Weak $\e$-Preservation]
\label{lemma:e-preservation-weak}
For all $\e, \e', \A$, if $\vdash \e : \A$ and $\e \leadsto \e'$, then $\vdash \e' : \A$.
\end{lemma}

This lemma also requires a linear version of substitution rule:
\[
\infer[\mbox{$\e$-Subst}]{\Gamma; \Delta_1, \Delta_2 \vdash \e[x \mapsto \e'] : \A}{
    \Gamma; \Delta_1, x : \A' \vdash \e : \A
    &
    \Gamma; \Delta_2 \vdash \e' : \A'
}
\]

The only tricky part for the proof of this lemma was an extra assumption of $x_1 \ne x_2$ for the following relation:
\[
\mathtt{let}\ (x_1,x_2)=(\e_1,\e_2)\ \mathtt{in}\ \e \leadsto \e[x_1 \mapsto \e_1, x_2 \mapsto \e_2].
\]

As the typing rules used in the proof of Lemma \ref{lemma:e-preservation-weak} were independent of $\Gamma$, a slightly stronger version of this lemma is also similarly provable:

\begin{lemma}[Slightly Stronger $\e$-Preservation]
\label{lemma:e-preservation-stronger}
For all $\Gamma, \e, \e', \A$, if $\Gamma; \emptyset \vdash \e : \A$ and $\e \leadsto \e'$, then $\Gamma; \emptyset \vdash \e' : \A$.
\end{lemma}